---
title: "Assignment 1"
author: "Pruthvi(CE16B034)"
date: "24/01/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Question 1
Create a matrix ‘M’ with 6 row and 6 column (with the column and row names).

```{r}
# Creating a matrix M with elements as the numbers from 1 to 36.
M = matrix(1:36,nrow=6,byrow=TRUE)
# Now we name the rows and columns of the matrix as row1, col1, and so on.
rownames(M) = c('row1','row2','row3','row4','row5','row6')
colnames(M) = c('col1','col2','col3','col4','col5','col6')
# display M
M
```

## (a)
Create a vector ‘V’ containing the diagonal elements of the above matrix ‘M’. (Don’t write manually!)

```{r}
# We use the diag function
V = diag(M)
# display the vector
V    
```

## (b)
Create a matrix ‘M1’ containing the first 3 rows and last 3 columns of ‘M’. (Don’t write manually!)

```{r}
# First 3 rows of M1 are first 3 rows of M and last 3 rows of M1 are last 3 columns of M
M1 = matrix(c(M[1,],M[2,],M[3,],M[,4],M[,5],M[,6]),nrow=6,byrow=TRUE)
# dislplay the required matrix
M1
```

## (c)
Convert the matrix ‘M’ into data-frame.

```{r}
# To convert the matrix 'M' into data-frame we use the 'as.data.frame()' method
as.data.frame(M)
```

# Question 2
Import the data file ‘parusater.txt’ from the web as:
Data <- read.table(“http://www.oikostat.ch/data/parusater.txt”, header=TRUE)

```{r}
# we use read.table to get the required data
Data<-read.table("http://www.oikostat.ch/data/parusater.txt",header = TRUE)
# display the read data
Data
```

## (a)
Extract the column names

```{r}
# We call the colnames function on 'Data' 
CName = colnames(Data)
# Display the extracted column names
CName
```

## (b)
Calculate the number of rows and columns of “Data”

```{r}
# nrows and ncols function can be directly used to get number of rows and columns directly
# We use cat function to display the string and variable output directly
cat("No of rows: ",nrow(Data))
cat("No of cols: ",ncol(Data))
```

## (c)
Extract the “age” from “Data” and store in a variable ‘Age’. Find the maximum and minimum value of ‘Age’

```{r}
# We use $ to extract the age as a vector
Age = Data$age
# Displays the extracted vector
Age
# Minimum
min(Age)
# Maximum
max(Age)
```

## (d)
Remove all the rows with the value of sex = 2 and store it in a new variable ‘NewData’
```{r}
# We are now deciding whether a particular row is present or not by comparing 'sex' column 
NewData = Data[Data$sex != 2,]
# Display Newdata
NewData
```

# Question 3
Create a matrix ‘Mat’ by rows instead of by columns of order 6 and some entries should be non-integer

```{r}
# We use seq with by = 0.5 to generate some non-integer values
Mat = matrix(seq(1,18.5,by=0.5),nrow=6,byrow=T)
# Display the matrix (6 * 6)
Mat
```
## (a)
Find the transpose of the matrix

```{r}
# We use 't()' function to generate transpose of the matrix
t(Mat)
```

## (b)
Calculate the determinant of the matrix

```{r}
# We use 'det()' function to find the determinant of the matrix
det(Mat)
```

## (c)
If the matrix is non-singular, calculate its inverse

```{r}
# Here the matrix is singular since the det(Mat) = 0, hence we can't find it's inverse
# However if the Matrix were non-singular we could've used solve(Mat) to find it's inverse
# solve(Mat)
```

## (d)
Assign 0 for non-integer entries. (Don’t assign 0 manually!)

```{r}
# is.integer() cannot be used here because R is treating all these values as doubles
# Instead we check for x != floor(x) , which means it is not integer
# Then we assign all such values to 0
Mat[Mat != floor(Mat)] = 0
# Display the modified Matrix
Mat
```

# Question 4
Create a data-frame ‘Cricket’ containing atleast 10 International cricket players details of ODI matches. The columns name are as follows: Name, Total Score, Highest Score, Century, Half-century, Wicket, Six, Four, Average, Strike Rate

```{r}
# vec1 contains the list of all cricket players
# All of the data below has been taken from https://en.wikipedia.org
vec1 = c("Sachin Tendulkar","Kumar Sangakkara","Ricky Ponting","Sanath Jayasuriya","Mahela Jayawardene","Inzamam ul-Haq","Jacques Kallis","Sourav Ganguly","Rahul Dravid","Virat Kohli")
# vec2 contains their total scores
vec2 = c(18426,14234,13704,13430,12650,11739,11739,11363,10889,10430)
# vec3 contains their highest scores across all cricket matches
vec3  = c(248,319,257,189,374,329,224,239,270,243)
# vec4 contains their Number of century's
vec4 = c(49,25,30,28,19,10,17,22,12,39)
# vec5 contains their Number of half-century's
vec5 = c(96,93,82,68,77,83,86,72,83,48)
# vec6 contains the number of wickets
vec6 = c(46,0,5,98,6,0,292,32,1,0)
# vec7 contains the number of 6's in Tests
vec7 = c(69,51,73,59,61,48,97,57,21,731)
# vec8 contains the number of 4's in Tests
vec8 = c(2058,1491,1508,910,1388,1105,1485,900,1654,19)
# vec9 contains the Average
vec9 = c(44.83,41.98,42.03,32.36,33.01,39.52,44.36,41.02,39.16,59.10)
# vec10 contains the Strike Rate
vec10 = c(86.23,78.88,80.39,91.20,78.67,74.24,72.89,73.70,71.24,92.65)
Cricket = data.frame(Name=vec1,TotalScore=vec2,HighestScore=vec3,Century=vec4,HalfCentury=vec5,Wicket=vec6,Six=vec7,Four=vec8,Average=vec9,StrikeRate=vec10,stringsAsFactors=F)
Cricket
# Summary
summary(Cricket)
```

