---
title: "Assingment 8"
author: "Pruthvi(CE16B034)"
date: "4/14/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Question 1
The data represent the diameter of pencils manufactured by a factory.
```{r}
# Using the xlsx library to read the given file.
library(xlsx)
# Using read.xlsx function to read sheet 1 of data
xl <- read.xlsx("QCC.xlsx",1)
xl
```

## (a) Calculate the control limits for X¯ chart and construct the chart.
```{r}
# constructing range, mean and sd values
xl <- cbind(xl,apply(xl,1,mean),t(diff(apply(xl[,1:4],1,range))),apply(xl,1,sd))
colnames(xl) <- c('x1','x2','x3','x4','x.bar','r','s')
xl
library(qcc)
batch = rep(1:30,each=4)
V <- qcc.groups(c(xl[,'x1'],xl[,'x2'],xl[,'x3'],xl[,'x4']),batch)
V
# X bar chart
q1 <- qcc(V,type = 'xbar')
print('From above graph information we see that,\nLCL = 63.17826\nUCL = 64.95315')
```

## (b) Determine the control limits for R and S charts and draw the charts.
```{r}
# R chart
q2 <- qcc(V,type='R')
print('From above graph information we see that,\nLCL = 0\nUCL = 2.779735')
# S chart
q3 <- qcc(V,type='S')
print('From above graph information we see that,\nLCL = 0\nUCL = 1.360366')
```

## (c) Is the process under control? If the answer is No, eliminate out of control points and construct revised control limits and control charts.
```{r}
# find out all the points violating the control process
shewhart.rules(q1)
shewhart.rules(q2)
shewhart.rules(q3)
out = c(11,26,14,29,7,18,19,20,21,22,23,24,25,8,9,10,16,3,17,27)
V2 = V[-out,]
V2
# Redraw the x bar plot, with revised control limits
q11=qcc(V2,type = "xbar")
# Redraw the R plot, with revised control limits
q22=qcc(V2,type = "R")
# Redraw the S plot, with revised control limits
q33=qcc(V2,type = "S")
```

## (d) If the specification limits are 64±0.05, Estimate Cp and Cpk, interpret these ratios.
```{r}
# get Cp and Cpk values
process.capability(q1,spec.limits=c(63.5,64.5))
print('From above graph and information,we have\nCp = 0.282\nCpk = 0.245')
print('Since Cp < 1, then the process is bound to produce rejections even when the mean is set on target.')
print('Here since Cpk is low, we have to consider ways to improve our process such as reducing its variance or shifting its location')
print('Here since Cp and Cpk are approximately equal, the process is centered between the specification limits')
```

# Question 2
The data represent the number of dictation mistakes in every 100 words.
```{r}
# Read sheet 2 of excel sheet
xm <- read.xlsx("QCC.xlsx",2)
xm
```

## (a) Which control chart is good for this type of data?
```{r}
print('For this type of data NP control Chart is used because in NP control chart we see the numbner of defectives in samples of given sizes')
```

## (b) Find the control limits and plot the chart.
```{r}
# Plot the xm type chart
q4 = qcc(xm,type="np",sizes = 100)
```

## (c) Is the process in control? If not, eliminate out of control points. Revise the control limits.
```{r}
# find the points which violate the control process
shewhart.rules(q4)
out = c(17)
xm2 = xm[-out,]
# Redraw the np plot plot with revised control limits.
q5 = qcc(xm2,type="np",sizes = 100)
```

