---
title: "Assignment3"
author: "Pruthvi(CE16B034)"
date: "10 February 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# Loading the required 'prob' library
library(prob)
```

# Question 1
A card is drawn at random from each of two well-shuffled packs of cards. What is the probability that at least one of them is the king of diamond?
```{r}
L <- cards()
# We need 2 samples with ordering and with replacement
M <- urnsamples(L,2,replace = T,ordered = T)
# Assigns equal probability to the entire sample space
N <- probspace(M)
# Prob that any one of them is King and Diamond
Prob(N,any(rank == 'K' & suit == 'Diamond'))
# Verification:
# Prob that atleast one of them is King of Diamond = 
# 1 - Probability that none of them are King of Diamond = 
1 - (51/52) ** 2
```

# Question 2
A die is rolled 4 times in succession. Find the probability of obtaining 3 at least once.
```{r}
# We need a sample space such that the die has benn rolled 4 times
A <- rolldie(4,makespace = T)
# displays the first 6 rows of the sample space
head(A)
# Prob that alleast one of them is 3
Prob(A,X1 == 3 | X2 == 3 | X3 == 3 | X4 == 3)
# Verification
# Prob of obtaining 3 atleast once = 
# 1 - Prob of not obtaining 3 on all four dice = 
1 - (5/6) ** 4
```

# Question 3
An urn contains three balls numbered 1,2 and 3. Two balls are drawn successively, the first ball drawn being replaced. Find the probability that the sum of the two numbers is 4.
```{r}
# We need two samples numbered from 1 to 3 with replacement and with order
C <- urnsamples(1:3,2,replace = T,ordered = T)
# Assign equal prob to the entire sample space
D <- probspace(C)
# Displays the first 6 rows of the sample space
head(C)
# Prob such that sum == 4
Prob(D,(X1 + X2) == 4)
# Verification
# All the possiblities such that sum == 4 are :
# (1,3) (3,1) (2,2) => no of possibilites = 3
# so required prob = 3 * (Prob of one possibility) = 
3 * (1/9)
```

# Question 4
A six-sided die is loaded in a way that each even face is twice as likely as each odd face. All even faces are equally likely, as are all odd faces. Construct a probabilistic model for a single roll of this die and find the probability that the outcome is less than 3.
```{r}
# We give use probspace function to give our custom probability
# It can be easily verified that :
# prob of odd num = 1/9 and prob of even num = 2/9
B <- probspace(rolldie(1),probs=c(1/9,2/9,1/9,2/9,1/9,2/9))
# Display the first 6 rows of the sample space
head(B)
# Prob that outcome less than 3
Prob(B,X1<3)
# Verification
# Prob that outcome less than 3 means only X1 = 1 and X1 = 2 are the possibilities
# Prob = 
1/9 + 2/9
```

# Question 5
We roll two fair 6-sided dice. Each one of the 36 possible outcomes is assumed to be equally likely.
```{r}
# roll the die 2 times
X <- rolldie(2,makespace = T)
# display the first 6 rows of the sample space
head(X)
```

## (a) 
Find the probability that doubles are rolled (both dice have the same faces up).
```{r}
# prob that both sides are equal
Prob(X,X1 == X2)
# Verification
# (1,1) , (2,2) ,...... until (6,6) are the possibilities
# => 6 possibilities , so prob =
6 * (1/36)
```

## (b)
Given that the roll results in a sum of 4 or less, find the conditional probability that doubles are rolled.
```{r}
# use subset function to define the respective events
eventX = subset(X,(X1 + X2) <= 4)
eventY = subset(X, X1 == X2)
# Prob of Y event given X event
Prob(eventY,given=eventX)
# Verification
# prob(X,given Y) = prob(X and Y) / prob(Y)
# = prob(roll is a double and sum <= 4) / prob(sum <= 4)
# = prob( (1,1) , (2,2) ) / prob( (1,3) , (3,1) , (2,2) , (1,1) , (1,2) , (2,1) ) = 
 2/6
```

## (c)
Find the probability that at least one die roll is a 6.
```{r}
# either first or second die is a 6
Prob(X,(X1 == 6) | (X2 == 6))
# verification
# 1 - prob(none of them is a 6) = 
1 - (5/6) ** 2
```

## (d)
Given that the two dice land on different numbers, find the conditional probability that at least one die roll is a 6.
```{r}
# numbers on two dice are not equal
eventX = subset(X,X1 != X2)
# either of the numbers on the dies is a six
eventY = subset(X,(X1 == 6) | (X2 == 6))
# prob of Y event given X
Prob(eventY,given=eventX)
# verification
# prob(X,given Y) = prob(X and Y) / prob(Y)
# = prob(atleast one die = 6 and both numbers are different) / prob(both numbers are diff)
# prob(both numbers are diff) = 30/36
# prob(atleast one die = 6 and both numbers are different) = 10/36
# So required prob = 
10/30
```

# Question 6
Consider two urns, the first with 5 red balls and 3 green balls, and the second with 2 red balls and 6 green balls. Your friend randomly selects one ball from the first urn and transfers it to the second urn, without disclosing the color of the ball. You select one ball from the second urn. What is the probability that the selected ball is red?
```{r}
# prob of red ball being selected = 5/8
# prob of green ball being selected = 3/8
A <- rep(c("red","green"),times=c(3,6))
B <- rep(c("red","green"),times=c(2,7))
# with prob 5/8, the second urn will have 3 red and 6 green and hence be 'C'
C <- probspace(urnsamples(A,1))
# with prob 3/8, the second urn will have 2 red and 7 green and hence be 'D'
D <- probspace(urnsamples(B,1))
# required prob = 
5/8 * Prob(C,isrep(C,"red",1)) + 3/8 * Prob(D,isrep(D,"red",1))
# Verification
# We use the total probability theorem, required prob = 
5/8 * 3/9 + 3/8 * 2/9
```
